Pre-processing high frequency animal tracking data

### 2020-10-05 03_smoothing_tracks.Rmd
some code edits

```{r}
# prepare data to plot
# make list of data to plot
data_plot <- append(list(data, data_errors), list_of_smooths)

# add offset
data_plot <- mapply(FUN = function(df, offset) {
  df[, x := x + offset]
}, data_plot, seq(from = 0, to = 1, length.out = length(data_plot)),
SIMPLIFY = FALSE)

# bind rows
data_plot <- rbindlist(data_plot, fill = TRUE)

plot_figure <- ggplot()+
  # geom_point(data = data_plot,
  #            aes(x, y),
  #            col = "darkgrey", 
  #            alpha = 1,
  #            shape = 1)+
  geom_path(data = data_plot,
            aes(x, y, 
                col = window_size,
                group = window_size),
            alpha = 1,
            show.legend = FALSE)+
  annotate(x = 0.6 + seq(0, 1, length.out = 6),
           y = 1.05,
           label = glue::glue('({letters[seq(6)]})'),
           geom = "text",
           fontface = "bold")+
  scale_colour_viridis_c(option = "D",
                         na.value = "red")+
  coord_equal() +
  theme_test()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())

plot_figure

# save figure
ggsave(plot_figure, filename = "figures/fig_median_smooth.png",
       width = 185, height = 185 * 2 / 3, units = "mm")
```

@@

```{r}
# choose the data
data_agg <- copy(data)

list_of_agg <- lapply(c(10, 30, 120), function(z) {
  data_agg <- atl_thin_data(data = data_agg,
                            interval = z,
                            method = "aggregate")
  data_agg[, interval := z]

})
```

```{r}
# plot
data_plot <- append(list(data[, `:=`(SD = 0, interval = NA)]), list_of_agg)

# add offset
data_plot <- mapply(FUN = function(df, offset) {
  df[, x := x + offset]
}, data_plot, seq(from = 0, to = 1, length.out = length(data_plot)),
SIMPLIFY = FALSE)

data_plot <- rbindlist(data_plot, fill = TRUE)

plot_figure <- ggplot(data = data_plot)+
  geom_point(aes(x, y, size = SD),
             colour = "darkblue",
             alpha = 0.5,
             shape = 16, show.legend = F)+
  geom_path(aes(x, y,
                group = interval,
                col = interval),
            col = "red",
            show.legend = FALSE)+
  annotate(x = 0.6 + seq(0, 1, length.out = 4),
           y = 1.05,
           label = glue::glue('({letters[seq(4)]})'),
           geom = "text",
           fontface = "bold")+
  coord_equal()+
  theme_test()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())

plot_figure

# save figure
ggsave(plot_figure, filename = "figures/fig_aggregate_data.png",
       width = 185, height = 90, units = "mm")
```

>>>>

### Robustness to large gaps

```{r}
# make a single list
data_plot <- append(list(data_errors), list_of_smooths)

# now prep gap
fig_smooth_1d <- 
  ggplot()+
  geom_path(data = data[1:2000,],
            aes(time, y),
            size = 0.2)+
  geom_point(data = rbindlist(data_plot),
             aes(time, y,
                 col = factor(window_size)),
             alpha = 0.5,
             size = 1,
             show.legend = FALSE,
             shape = 16)+
  coord_cartesian(xlim = c(5000, 7000),
                  ylim = c(0.7, NA),
                  expand = F)+
  scale_colour_brewer(palette = "Spectral")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank())+
  labs(x = "position index", y = "Y coordinate")
```

### Plot results

```{r}
# prepare data to plot
# make list of data to plot
figure_median_smooth <- 
  mapply(function(df, col) {
    ggplot()+
      geom_point(data = df,
                 aes(x, y),
                 size = 0.5,
                 alpha = 0.5,
                 show.legend = FALSE,
                 col = col,
                 shape = 1)+
      geom_path(data = data,
                aes(x, y),
                col = "grey20")+
      coord_cartesian(expand = T,
                  ylim = c(0.6, 0.8)) +
      ggthemes::theme_map()
}, data_plot, c(pal[2], rep(
  RColorBrewer::brewer.pal(3, "Greens")[2],
  4)),
SIMPLIFY = F)

# make basic plot with rectangle
fig_bounds <- 
  ggplot()+
  geom_path(data = data,
            aes(x, y))+
  geom_hline(yintercept = c(0.6, 0.8),
             col = "grey20",
             lty = 2, lwd = 0.2)+
  coord_equal(expand = F) +
  ggthemes::theme_map()

# append to first figure
figure_median_smooth[[1]] <-
  figure_median_smooth[[1]]+
  annotation_custom(grob = ggplotGrob(fig_bounds),
                    ymin = 0.6, ymax = 0.8,
                    xmin = 0.5, xmax = 0.7)

# wrap plots
fig_median_smooth <- wrap_plots(append(figure_median_smooth, 
                                       list(fig_smooth_1d)), 
                                       ncol = 3)+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

fig_median_smooth

# save figure
ggsave(fig_median_smooth, filename = "figures/fig_median_smooth.png",
       width = 225, height = 225 / 1.77, units = "mm")
```

@@

```{r}
# choose the data
data_agg <- copy(list_of_smooths[[3]])

# get list of aggregated data
list_of_agg <- lapply(c(3, 10, 30, 120), function(z) {
  data_return <- atl_thin_data(data = data_agg,
                            interval = z,
                            method = "aggregate")
  data_return[, interval := z]
  return(data_return)
})
```

```{r}
# make figure
figure_aggregate <- 
  mapply(FUN = function(df) {
  ggplot(data = df)+
      geom_point(data = list_of_smooths[[3]],
                 aes(x, y),
                 col = pal[3],
                 size = 0.3, alpha = 0.5)+
      geom_point(aes(x, y, size = SD),
                 colour = pal[4],
                 alpha = 1,
                 stroke = 1,
                 shape = 4, show.legend = F)+
      geom_path(aes(x, y),
                 colour = pal[4],
                 alpha = 1)+
      geom_path(data = data,
                lwd = 0.2,
                aes(x, y),
                col = "grey20")+
      coord_equal(ylim = c(0.6, 0.8))+
      ggthemes::theme_few()+
      theme(axis.title = element_blank(),
            axis.text = element_blank())
}, list_of_agg[2:3],
SIMPLIFY = FALSE)

figure_aggregate <- wrap_plots(figure_aggregate)

figure_aggregate
# save figure
ggsave(plot_figure, filename = "figures/fig_aggregate_data.png",
       width = 185, height = 90, units = "mm")
```

### Aggregation before correction, a warning

```{r}
# read data with errors
data_errors <- fread("data/data_errors.csv")

# aggregate before correction
list_of_agg <- lapply(c(3, 10, 30, 120), function(z) {
  data_return <- atl_thin_data(data = data_errors,
                            interval = z,
                            method = "aggregate")
  data_return[, interval := z]
  data_return[, speed := atl_get_speed(data_return)]
  return(data_return)
})

# get real speed
data[, speed := atl_get_speed(data)]

### plot figures
fig_agg_data <- 
  lapply(list_of_agg, function(df) {
    ggplot(df)+
      geom_path(aes(x,y), 
                # col = factor(interval), 
                # group = interval),
                lwd = 0.2)+
      geom_point(aes(x,y, 
                     size = SD,
                     shape = speed < quantile(data$speed, 0.95, na.rm = T),
                     col = speed < quantile(data$speed, 0.95, na.rm = T),
                     group = interval),
                 stroke = ifelse(df$speed < quantile(data$speed, 0.95, na.rm = T),
                                 1, 1),
                 show.legend = F)+
      geom_path(data = data,
                aes(x, y),
                lwd = 0.1)+
      scale_shape_manual(values = c(2, 4))+
      scale_colour_manual(values = pal[c(1, 2)])+
      coord_cartesian(ylim = c(0.6, NA))+
      ggthemes::theme_map()  
  })
```

```{r}
# now plot hist of speed
data_agg <- rbindlist(list_of_agg)

# show boxplot of speed
fig_agg_speed <- ggplot(data_agg)+
  geom_jitter(aes(factor(interval), speed),
              col = "grey",
              shape = 4, alpha = 0.5)+
  geom_boxplot(aes(factor(interval), speed),
               fill = NA,
               show.legend = F, 
               width = 0.3, 
               outlier.size = 0.2)+
  geom_hline(yintercept =
               c(mean(data$speed, na.rm = T),
                 quantile(data$speed, na.rm = T,
                          probs = c(0.95))
               ),
             lty = c(1, 2))+
  scale_y_log10(label = scales::comma)+
  ggthemes::theme_few()+
  theme(axis.text.y = element_blank())+
  labs(x = "interval (s)",
       y = "speed")
```

### Combine with aggregations

```{r}
# make combined figure
fig_aggregate <- 
  wrap_plots(append(fig_agg_data[2:3], list(fig_agg_speed)),
           design = "AAABBBCC")+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(fig_aggregate,
       filename = "figures/fig_aggregate_data.png",
       width = 170, height = 170 / 1.67, units = "mm")
```

### 2020-10-05 11_residence_patch.Rmd
some code edits

```{r prep_libs_02_01}
library(data.table)
library(atlastools)
library(ggplot2)
```

@@

# make residence patch
patch <- atl_res_patch(data, 
                       buffer_radius = 0.0001, 
                       lim_spat_indep = 0.02, 
                       lim_time_indep = 3)
# get spatial representation
patch_sf <- atl_patch_summary(patch_data = patch, 
                              which_data = "spatial", 
                              buffer_radius = 0.01)

@@

```{r}
plot_patches <- ggplot()+
  geom_sf(data = patch_sf,
          colour = "NA",
          aes(fill = duration),
          show.legend = FALSE)+
  geom_point(data = data,
             aes(x, y),
             size = 0.1,
             alpha = 0.1)+
  geom_point(data = patch_summary,
             aes(x_median + 0.5, y_median),
             size = 5,
             shape = 21,
             fill = "grey")+
  annotate(geom = "text",
           x = patch_summary$x_median + 0.5, 
           y = patch_summary$y_median + 0.02,
           label = patch_summary$patch,
           col = "red")+
  annotate(geom = "text",
           x = patch_summary$x_median, 
           y = patch_summary$y_median + 0.02,
           label = patch_summary$patch,
           col = "red")+
  annotate(geom = "text",
           x = c(0.5, 1.1),
           y = 1,
           label = sprintf("(%s)", c("a", "b")),
           fontface = "bold")+
  geom_path(data = patch_summary,
            aes(x_median + 0.5, y_median),
            colour = "darkblue",
            arrow = arrow(angle = 10, 
                          type = "closed"))+
  scale_fill_viridis_c()+
  theme_void()

@@

# save figure
ggsave(plot_patches, 
       filename = "figures/fig_residence_patch.png",
       width = 185, height = 185 * 2 / 3, units = "mm")
```

>>>>

```{r prep_libs_02_01}
library(data.table)
library(atlastools)
library(ggplot2)

# prepare a palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

@@

# make residence patch
patch <- atl_res_patch(data, 
                       buffer_radius = 0.0001, 
                       lim_spat_indep = 0.02, 
                       lim_time_indep = 3)

# get spatial representation
patch_sf <- atl_patch_summary(patch_data = patch, 
                              which_data = "spatial", 
                              buffer_radius = 0.01)

@@

```{r}
plot_patches <-
  ggplot()+
  geom_sf(data = patch_sf,
          colour = "grey20",
          fill = "grey90",
          show.legend = FALSE)+
  geom_point(data = patch_summary,
             aes(x_median + 0.5, 
                 y_median,
                 size = duration),
             # size = 8,
             shape = 2,
             col = pal[2],
             stroke = 2,
             show.legend = F)+
  geom_path(data = data,
             aes(x, y),
            size = 0.3)+
  geom_curve(data = patch_summary,
            aes(x = x_end + 0.5, y = y_end,
                xend = shift(patch_summary$x_start, 
                             type = "lag") + 0.5,
                yend = shift(patch_summary$y_start, 
                             type = "lag")),
            col = "grey30",
            size = 0.3,
            ncp = 2,
            curvature = -0.4,
            arrow = arrow(angle = 20,
                          length = unit(2, units = "mm"), 
                          type = "closed", ends = "first"))+
  annotate(geom = "text",
           x = c(0.5, 1.1),
           y = 1,
           label = sprintf("(%s)", c("a", "b")),
           fontface = "bold")+
  scale_colour_distiller(palette = "YlGnBu",
                         trans = "log10")+
  coord_sf()+
  ggthemes::theme_map()+
  theme(legend.position = "right")

@@

# save figure
ggsave(plot_patches, 
       filename = "figures/fig_residence_patch.png",
       width = 170, height = 100, units = "mm")
```
